---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r libraries, echo = FALSE, message=F, warning=FALSE}
library(lubridate)
library(vkR)
library(network)
library(sna)
library(igraph)
library(tibble)
#library(statnet)
# detach(package:sna)
# detach(package:network) 
# detach(package:igraph) 

```

There are two communities in VK dedicated to meme in cognitive psychology field: ["Cognitive partimaker"](https://vk.com/cogparty) and ["Tyagoty bytiya poznayuschego"](https://vk.com/subro_sa).
And these communities are in the constant conflict with each other due to different attitudes in psychology.    
But I suppose that a lot of people are in the both of them. I am interested to check is it true or not and who are that people who are only in the one community.


# Getting token 
```{r}
token <- '7039f8fcd4af1aa3f126ff79e9d1224d7453dc881605c01191b2215b0b8350656f943ef02e51520ab2b9c'
setAccessToken(token)
areFriends(1,2)
```




# Tyagoty bytiya

##Parsing data
```{r}
tyagoty_id <- 155160218
#getGroupsMembersExecute(tygoty_id)
usersTyagoty <- getGroupsMembersExecute(tyagoty_id, fields = 'first_name, last_name, sex, education')
head(usersTyagoty)

```

```{r}
usersTyagoty$fullname <- paste0(usersTyagoty$first_name, usersTyagoty$last_name)
usersTyagoty <- add_column(usersTyagoty, fullname = paste0(usersTyagoty$first_name, ' ', usersTyagoty$last_name), 
                           .after = "id")
usersTyagoty_adj <- getArbitraryNetwork(usersTyagoty$id, format = 'adjmatrix')

usersTyagoty_edge <- getArbitraryNetwork(usersTyagoty$id, format = "edgelist")
View(usersTyagoty_adj)
usersTyagoty_net_igraph <- graph.adjacency(usersTyagoty_adj, mode=c("undirected"), weighted=NULL, diag=FALSE)
usersTyagoty_net_igraph2 <- graph_from_data_frame(usersTyagoty_edge, directed = F, vertices = usersTyagoty)
```

##Add atributes to the graph 
```{r}
list.vertex.attributes(usersTyagoty_net_igraph2)
head(vertex_attr(usersTyagoty_net_igraph, 'name'))
tp <- merge(as_data_frame(usersTyagoty_net_igraph, 'vertices'), usersTyagoty_table, by.x = 'name', by.y = "id")
usersParty_net_igraph <- set_vertex_attr(usersParty_net_igraph, 'surname', value = tp$last_name)

tp$last_name
length(tp$last_name)

length(as_data_frame(usersTyagoty_net_igraph, 'vertices')$name)

list.vertex.attributes(usersTyagoty_net_igraph)
```

##Centrality measures

```{r}
tyagoty_degree <- degree(usersTyagoty_net_igraph2, mode = 'all')
class(tyagoty_degree)
head(tyagoty_degree)
tyagoty_degree
tyagoty_betweenness <- betweenness(usersTyagoty_net_igraph2, directed = F, weights = NULL)
tyagoty_closeness <- closeness(usersTyagoty_net_igraph2, mode = 'all', weights = NULL)
tyagoty_eigen <- eigen_centrality(usersTyagoty_net_igraph2, directed = F, weights = NULL)

tyagity_measures <-  cbind(usersTyagoty, tyagoty_degree, tyagoty_betweenness, tyagoty_closeness, tyagoty_eigen)
merge
head(usersTyagoty)
```


```{r}
plot(usersTyagoty_net_igraph2,
     vertex.size = 1, 
     edge.arrow.size = 0.1, 
     vertex.label.cex = condSize,
     # vertex.color = colors_spectr[9],
     # edge.color=colors_greys[3], 
     vertex.shape = 'circle', 
     vertex.label = usersTyagoty$fullname,
     vertex.label.dist = 1, 
     vertex.label.degree = 5,
     vertex.label.font = 1,
     vertex.label.color = condColor, 
     lty=2)

```






