---
title: "Seminar 2"
author: "Elena Rybina 'MKH-161'"
date: "1/22/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = FALSE, message=F, warning=FALSE}
library('network')
library(foreign)
```
## Reading network data in R  

```{r matrix of edges, , fig.align='center',include=TRUE}
# Let's have 15 nodes (you can choose any number).
num_nodes <- 15

# Next, generate the matrix using the build-in "matrix" command in R 
my_matrix<-matrix(round(runif(num_nodes*num_nodes)), # edge values
                  nrow = num_nodes, #nrow must be same as ncol
                  ncol = num_nodes)

# Next, let's make sure there are no self-referencing loops
# (meaning, the node is not connected to itself)

diag(my_matrix) <- 0

# We can check dimensions of the new object: 
dim(my_matrix)

class(my_matrix)

sum(is.na(my_matrix))
```


```{r  as.network, , fig.align='center', include=TRUE}

##?as.network

#matrix to network
my_network<-as.network(x = my_matrix, # the network object
                       directed = TRUE, # specify whether the network is directed 
                       loops = FALSE, # do we allow self ties (should not allow them) 
                       matrix.type = "adjacency" # the type of input  adjencecy -- symmetric matrix
)


#kind of desctiptive statistics
network.size(my_network)
summary(my_network)


# View your network:
par(mar=c(1,1,1,1)) # get rid of the large margins and plot the data:
plot(my_network)
```


## Importing other data formats


```{r pajek format, fig.align='center'}
drugpaj <- read.paj('drugnet2.paj')

names(drugpaj) # objects in drugpaj
drugpaj$networks
names(drugpaj$networks)
names(drugpaj$partitions)

drug <- drugpaj$networks[[1]] # extract network 
class(drug)
plot(drug)

```



***Assignment question: what looks strange about this network? Why?***

This network has a lot of nodes that doesn't connect with any other node. It seems strange that people in the one communitiy and one network could be completely disconnected. ??especially for RSD design??

```{r, fig.align='center'}
network.size(drug) # how many nodes?
network.edgecount(drug) # how many edges?
network.dyadcount(drug) # how many dyads?
network.edgecount(drug)/network.dyadcount(drug)  #dencity of network
```


***Assignment question: What do the numbers above represent?***    

The first number `r network.size(drug)` is amount of nodes (how many people are).  
The second one `r network.edgecount(drug)` is amount of edges or ties or connections between nodes.  
The third one 85556 is amount of all possible connections between each two nodes in this network.  


***Assignment task: For the network “drug” that we created and loaded with attributes, create several different network plots, adding gender and ethnicity to the graph as node attributes. Consider using a variety of colors to make your plot more informative.***

```{r pajek attributes, fig.align='center'}
names(drugpaj$partitions)

#-----gender--------
gender <- drugpaj$partitions[[1]]
gender
table(gender)

colors <- ifelse(gender==2, 'red', ifelse(gender==1, 'navyblue', 'grey'))

female <- ifelse(gender == 2, 1, # recode
                 ifelse(gender == 1, 0, NA))

## set attributes
drug <-set.vertex.attribute(drug, 'female', value=c(female)) 
plot(drug, vertex.col=colors)


#----ethnicity----

ethnicity <- drugpaj$partitions[[2]] # extract ethnicity 
ethnicity
table(ethnicity) # how is it coded?
colors2 <- ifelse(ethnicity==1, 'red', ifelse(ethnicity==2, 'blue', ifelse(ethnicity==3, 'yellow', 'grey')))

plot(drug, vertex.col=colors2)

```


## Reading a native R data file


```{r inner format, fig.align='center'}
load('flo.Rdata')



flo.marriage <- as.network(as.matrix(flo.marriage), directed=FALSE)
flo.biz <- as.network(as.matrix(flo.biz), directed=FALSE)
flo.att[,1]
#flo.att[1,]

# Add attributes
flo.marriage

flo.att

set.vertex.attribute(flo.marriage, 'wealth', flo.att[,2]) 
set.vertex.attribute(flo.biz,'wealth', flo.att[,2])

# Simple plots:
par(mfrow=c(1,2))  # par -- paragraph
#first number -- number of columns, the second -- rows
par(mar=c(0,0,0,0)) #margins
plot(flo.marriage, displaylabels=TRUE)

```

```{r, fig.align='center'}
par(mar=c(0,0,0,0)) #margins
plot(flo.biz, displaylabels=TRUE)
```


```{r inner format attributes, , fig.align='center'}
#Plots with attributes

plot(flo.marriage,
     vertex.cex=(get.vertex.attribute(flo.marriage, 'wealth')/25 +.4),
     displaylabels=TRUE, label.cex=.5, label.pos=0, coord=plot(flo.marriage))
```


***Assignment task: Please plot the biz network with node attributes that you’ve set above.***

```{r inner format attributes 2, fig.align='center'}


plot(flo.biz,
     vertex.cex=(get.vertex.attribute(flo.biz, 'wealth')/25 +.4),
     displaylabels=TRUE, label.cex=.5, label.pos=0, coord=plot(flo.biz))

```


## Importing UCINet files directly into R

```{r UCINet format, fig.align='center', message=F}
# Read vertex labels with scan()
flo.names <- scan('padgett.dat', what='character', skip=4, nlines=16) # Read data with read.table()
flos <- read.table('padgett.dat', skip=41, col.names=flo.names)
# Read node attributes:
flo.att <- read.table('padgw.dat', col.names=c('WEALTH','NUM.PRIORS','NUM.TIES'), skip=25) 
flo.att
flo.att <-cbind(flo.names,flo.att)
# Command below provides first six rows of data, so that you can check your data without l 
head(flo.att)
 # Separate adjacency matrices
flo.marriage <-flos[1:16,] # subset of the first 16 colums is the marriage network 
dim(flo.marriage)
row.names(flo.marriage) <-flo.names # name
flo.biz <- flos[17:32,] # subset of the second 16 is the business network. 
row.names(flo.biz) <-flo.names # name
dim(flo.biz)
# Check the data by listing a couple of rows and colums from each network.
flo.marriage[1:2,1:2]
flo.marriage <- as.network(as.matrix(flo.marriage),directed=FALSE) 
flo.biz <- as.network(as.matrix(flo.biz),directed=FALSE)
## add attributes
set.vertex.attribute(flo.marriage, 'wealth', flo.att[,2]) 
set.vertex.attribute(flo.biz,'wealth', flo.att[,2])
```

***Assignment task: Using the code already shown, plot both of the new networks. Add attributes if you wish.***

``` {r, fig.align='center'}
#the same plots as previous plots
plot(flo.marriage,
     vertex.cex=(get.vertex.attribute(flo.marriage, 'wealth')/25 +.4),
     displaylabels=TRUE, label.cex=.5, label.pos=0, coord=plot(flo.marriage))
plot(flo.biz,
     vertex.cex=(get.vertex.attribute(flo.biz, 'wealth')/25 +.4),
     displaylabels=TRUE, label.cex=.5, label.pos=0, coord=plot(flo.biz))

```

## Saving Network Data in R

```{r}
# Save several objects in the same .Rdata file to load all at once
save(flo.marriage, flo.biz, file='floNets.Rdata')
# Save network as an edgelist in a .csv file
drug.edges <- as.matrix(drug, matrix.type='edgelist') 
write.csv(drug.edges, file='drugsEdgelist.csv',row.names=FALSE)
```

