---
title: "Seminar 9"
author: "Elena Rybina 'MKH 161'"
date: "3/21/2018"
output:
  pdf_document: default
  html_notebook: default
---
  
  
```{r libraries, echo = FALSE, message=F, warning=FALSE}
library(network)
library(sna)
library(RColorBrewer) 
library(NetData)
library(statnet)
library(coda)
library(ergm)
library(rgl)
library(stats)
```

#ERGM

##ERGM creation, continued


```{r}
data(studentnets.ergm173, package = "NetData")
class(edges)
head(edges)
dim(edges)
head(nodes)
dim(nodes)
```


To execute the ERGM, R requires continuous integer IDs: [1:n], where n is the total number of nodes in the ERGM. So, create node IDs acceptable to R and map these to the edges.
```{r}
#create vector of IDs
id <- seq(1,22,1) 
nodes1<-cbind(id, nodes) 
class(nodes1)
head(nodes1)
nodes1[,1:2]


#correspondance
edges2<-merge(nodes1[,1:2], edges, by.x = 'std_id', by.y='alter_id') 
head(edges2) #check the new edge dataset
dim(edges2)

names(edges2)[1]<-"alter_id"
names(edges2)[2]<-"alter_R_id"
edges3<- merge(nodes1[,1:2], edges2, by.x = "std_id", by.y="ego_id") #data with IDs we need
names(edges3)[1]<-"ego_id" 
names(edges3)[2]<-"ego_R_id" 
head(edges3)
```

##Creating the network and moving onto ERGM

```{r, fig.align='center'}
net<-network(edges3[,c("ego_R_id", "alter_R_id")]) #Dyads will be the unit of analysis 
# Assign edge-level attributes - dyad attributes
edges[,2]
set.edge.attribute(net, "ego_R_id", edges[,2])
set.edge.attribute(net, "alter_R_id", edges[,4])
# Assign node-level attributes to actors in "net"
net %v% "gender" <- nodes1[,3] 
net %v% "grade" <- nodes1[,4] 
net %v% "race" <- nodes1[,5]
net %v% "pci" <- nodes1[,6] 
plot(net) # Look at the network
```


```{r}
m1<-ergm(net ~ edges + mutual + nodematch("gender") + absdiff("pci"),burnin=15000,MCMCsamplesize=30000,verbose=FALSE)
#incomes are cored absolute difference

summary(m1)
```


```{r}
lapply(m1[1],exp) #this gets us a vector of exponents

exp(m1[1]$coef)

```

*******    

_**Assignment task:**_
*1. For the model above, explain the effect of absolute differences in pci on tie formation.* 
Rought exponential form of coefficient = 1.00011144 we can say that absolute differences in pci is almost the same as in the random graph, so, it doesn't affect tie formation.
*2. For the coefficients above, calculate the associated probabilities.*    
```{r}
#edges
exp(-2.316e+00)/(1+exp(-2.316e+00))
#mutual
exp(2.420e+00)/(1+exp(2.420e+00))
#nodematch.gender
exp(1.674e-02)/(1+exp(1.674e-02))
#absolute differences in pci
exp(1.114e-04)/(1+exp(1.114e-04))


```


Log-odds for edges is -2.316e+00, probability to form a tie is p = exp(-2.316e+00)/(1+exp(-2.316e+00)) = `r exp(-2.316e+00)/(1+exp(-2.316e+00))`.    
Log-odds for mutual is 2.420e+00, probability that tie will be mutual is p = exp(2.420e+00)/(1+exp(2.420e+00)) = `r exp(2.420e+00)/(1+exp(2.420e+00))`. 
Log-odds for nodematch.gender is 1.674e-02, probability for form tie with matching gender is p = exp(1.674e-02)/(1+exp(1.674e-02)) = `r exp(1.674e-02)/(1+exp(1.674e-02))`.    
Log-odds for absolute differences in pci is 1.114e-04, probability is p = exp(1.114e-04)/(1+exp(1.114e-04)) = `r exp(1.114e-04)/(1+exp(1.114e-04))`.   

*3. Re-run the model with only the significant elements present. For the obtained coefficients, calculate the associated probabilities. What do you observe?*   

```{r}
m2<-ergm(net ~ edges + mutual +  absdiff("pci"),burnin=15000,MCMCsamplesize=30000,verbose=FALSE)
summary(m2)
```
Log-odds for edges is -2.304e+00, probability to form a tie is p = exp(-2.304e+00 )/(1+exp(-2.304e+00 )) = `r exp(-2.304e+00 )/(1+exp(-2.304e+00 ))`.    
Log-odds for mutual is  2.410e+00, probability that tie will be mutual is p = exp( 2.410e+00)/(1+exp( 2.410e+00)) = `r exp( 2.410e+00)/(1+exp( 2.410e+00))`. 
Log-odds for absolute differences in pci is 1.106e-04, probability is p = exp(1.106e-04)/(1+exp(1.106e-04)) = `r exp(1.106e-04)/(1+exp(1.106e-04))`.     

*******   

##Model Diagnostics: troubleshooting and checking for model degeneracy    

```{r simulation}

m1.sim<-simulate(m1, nsim=10) # nsim is the number of simulated models
class(m1.sim) #check what we get
m1.sim
summary(m1.sim)


```

```{r, fig.align='center'}
m1.sim[[1]] #first simulated model
m1.sim[[2]] #second simulated model
# We can also look at them:
par(mfrow=c(1,3), mar=c(0,0,0,0)) 
plot(net) # original model 
plot(m1.sim[[1]]) 
plot(m1.sim[[2]])
```
******    


**Assignment task:** *For the model you created above (with only the significant coefficients) create a set of simulated models and plot a few against the original. What do you observe?*

```{r, fig.align='center'}
m2.sim<-simulate(m2, nsim=5) # nsim is the number of simulated models
m2.sim
summary(m2.sim)
m2.sim[[1]] 
m2.sim[[2]] 
par(mfrow=c(1,3), mar=c(0,0,0,0)) 
plot(net) # original model 
plot(m2.sim[[1]]) 
plot(m2.sim[[2]])
```


******    

##Goodness of fit, "gof"


```{r checking on data, fig.align='center'}
library(knitr)
m1.gof<-gof(m1~idegree)
m1.gof
# The code below is commented out, but if you want to see
# what we get with m1.gof object, check it:
names(m1.gof)
m1.gof$pval.ideg
kable(m1.gof$pval.ideg, caption="Goodness-of-fit for Indegree")

 par(mfrow=c(1,1)) 
 plot(m1.gof)
```

******    

_**Assignment task:**_ *For the model you created above (with only the significant coefficients) run a goodness- of-fit test with any network characteristic you choose, which is not already in the model. Generate a table with information and a plot. What do you observe?*


```{r, fig.align='center'}
m2.gof<-gof(m2~distance)
m2.gof
# The code below is commented out, but if you want to see
# what we get with m1.gof object, check it:
names(m2.gof)
m2.gof$pval.ideg
kable(m2.gof$pval.ideg, caption="Goodness-of-fit for Geodesic distances")

par(mfrow=c(1,1)) 
plot(m2.gof)
```

We can see from the gof grapth that our model fits geodesic deistances of the observable data very well.

*****   

##Markov Chain Monte Carlo, MCMC diagnostics

```{r, fig.align='center'}
par(mar=c(0,0,0,0)) 
mcmc.diagnostics(m1)
```

```{r, fig.align='center'}
data('faux.magnolia.high') 
magnolia <- faux.magnolia.high 
par(mar=c(0,0,0,0)) 
plot(magnolia, vertex.cex=.5)

```
```{r}
#fit <- ergm(magnolia~edges+triangle, control=control.ergm(seed=1))
```
```{r}
fit <- ergm(magnolia~edges+triangle, control=control.ergm(seed=1,MCMLE.maxit=3))
mcmc.diagnostics(fit)

```
```{r}
fit2 <- ergm(magnolia~edges+gwesp(0.5,fixed=T),control = control.ergm(seed=1))
mcmc.diagnostics(fit2)
```

```{r}
fit3 <- ergm(magnolia~edges+gwesp(0.25,fixed=T),control = control.ergm(seed=1), verbose=T)
mcmc.diagnostics(fit3)
```

```{r}
fit4 <- ergm(magnolia~edges+gwesp(0.25,fixed=T)+ nodematch('Grade')+nodematch('Race')+nodematch('Sex'),
control = control.ergm(seed=1,MCMC.samplesize=50000,MCMC.interval=1000), verbose=T)
mcmc.diagnostics(fit4)
```

